---
title: Characterizing Berlin's Districts Through Yelp Data
author: Gershom Tripp
date: '2018-12-14'
slug: characterizing-berlin-s-districts-through-yelp-data
categories: []
tags:
  - Berlin
  - Dataviz
  - Germany
  - Maps
  - R
  - Yelp
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

```{r load_packages, include=FALSE}
library(tidyverse)
library(ggmap)
```

```{r load_data, include=FALSE}
# load cleaned yelp data (found here: https://github.com/GershTri/yelp-r)
load("datasets/yelp/berlin_yelp.rda")
load("datasets/berlin_demographics/berlin_demo_sf_info.rda")
rm(pol_parties, info_district_df, sf_plz_berlin)
```

```{r reduce_restructure_yelp_df, include=FALSE}
# extract categories from categories.title.1
cat_titles1 <- yelp_df %>% 
    filter(!is.na(categories.title.1)) %>% 
    select(-categories.title) %>% 
    rename(categories.title = categories.title.1)

# extract categories from categories.title.2
cat_titles2 <- yelp_df %>% 
    filter(!is.na(categories.title.2)) %>% 
    select(-categories.title) %>% 
    rename(categories.title = categories.title.2)

# recombind above data frames with yelp_df and remove unneeded variables
yelp_df <- bind_rows(yelp_df, cat_titles1, cat_titles2) %>% 
    select(-c(alias, image_url, is_closed, url, starts_with("categories.alias"),
              starts_with("location.address"), starts_with("location.disp"),
              phone, display_phone, starts_with("categories.title."),
              location.country
              )
           )

rm(cat_titles1, cat_titles2)
```









## Introduction

***









<br><br>

## Getting the Data

***

The data were collected using the official Yelp [_Fusion_](https://www.yelp.de/developers/documentation/v3) API. Yelp's API isn't
really made for bulk downloading data, so there were three limits to 
contend with: 1. max 50 listings per request, 2. max 1000 listings per query (50 
listing-chunks x 20 offsets), and 3. max 5000 requests per day. Using district
names and PLZ (German ZIP codes) consistently failed because of 2., so I divided
up Berlin's coordinates into a grid of approx. 500x500-meter blocks and sent
requests for each block.

The above method worked well, except for one unforeseen problem: Longitudinal
coordinates do not run parallel to one another, but converge at the
poles and spread as they approach the equator, so the same longitudinal 
coordinates do not represent the same distances at different latitudes. The
upshot was that my grid-scan strategy broke down the more it went east and south.
Some sort of correction would be ideal, but I decided for a "good enough" approach
and simply shortened the longitudinal intervals and rescanned the affected parts.

This strategy, being more ax than scalpel, generates a lot of duplicates. After 
removing them I ended up with `r length(unique(yelp_df$id))` unique listings for 
Berlin and Potsdam, which is a small city just outside of Berlin. 











<br><br>

## A Brief Analysis of Yelp-Reviewed Businesses in Berlin

***

Before I get to the districts, I 

First, a glimpse of the data.

```{r show_glimpse}
# filter for berlin and potsdam
yelp_df <- yelp_df %>% 
    filter(location.state == "BE" | location.city %in% c("Berlin", "Potsdam"))

glimpse(yelp_df)
```

As I mentioned above, the number of unique listings for Berlin and Potsdam is 
just under 12,000. The 17,976 you see here is due the fact that some listings 
belong to multiple categories. I haven't been able to discern any kind of 
consistent hierarchy (i.e. supercategories and subcategories), so every 
A dot density plot can give some perspective on the distribution of Yelp-reviewed
businesses. The colors correspond to Berlin's 12 administrative districts and 
Potsdam, a town to the southwest of Berlin. The names of the districts aren't 
important at this point. Interesting is the disparity in the spatial density of
Yelp-reviewed businesses between districts.

```{r plot_dot_density_all_businesses, message=FALSE, warning=FALSE}
berlin_coord <- c(left = 12.9680, bottom = 52.3219, right = 13.8000, top = 52.6918)
berlin_map <- get_stamenmap(berlin_coord, maptype = "toner-background")

yelp_df[is.na(yelp_df$district), "district"] <- "Potsdam"

berlin_map %>% 
    ggmap(darken = .8) +
    geom_point(data = yelp_df %>%
                   select(id,
                          coordinates.latitude,
                          coordinates.longitude,
                          district
                          ) %>% 
                   distinct(),
               aes(coordinates.longitude,
                   coordinates.latitude,
                   color = district
                   ),
               size = 0.1,
               alpha = .3
               ) + 
    ggtitle("Distribution of All Yelp-Reviewed Businesses") + 
    theme(axis.text = element_blank(),
          axis.title = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none"
          )
```



```{r plot_30_categories_with_most_listings}
#
yelp_df %>% 
    group_by(categories.title) %>% 
    count() %>% 
    arrange(desc(n)) %>%
    head(30) %>% 
    rename(Category = categories.title, Count = n) %>% 
    ungroup() %>% 
    mutate(Category = factor(Category,
                             ordered = T,
                             levels = rev(.$Category)
                             )
           ) %>% 
    ggplot(aes(y = Category, x = Count)) +
    geom_point(shape = 21, fill = "lightblue", size = 2) + 
    scale_x_continuous(breaks = seq(100, 1700, 200)) +
    theme_minimal() +
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          axis.title.y = element_blank()
          ) + 
    ggtitle("Thirty Categories with the Most Yelp Listings")
```

```{r plot_30_categories_with_most_reviews_per_listing}
#
yelp_df %>% 
    group_by(categories.title) %>% 
    summarise(mean_review = mean(review_count, trim = .1)) %>% 
    arrange(desc(mean_review)) %>%
    head(30) %>% 
    rename(Category = categories.title) %>% 
    ungroup() %>% 
    mutate(Category = factor(Category,
                             ordered = T,
                             levels = rev(.$Category)
                             )
           ) %>% 
    ggplot(aes(y = Category, x = mean_review)) +
    geom_point(shape = 21, fill = "#F1C40F", size = 2) + 
    theme_minimal() +
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          axis.title.y = element_blank()
          ) + 
    ggtitle("Thirty Categories with Most Reviews (Mean) per Listing") +
    xlab("Reviews per Listing (Mean)")
```


```{r plot_30_businesses_with_most_reviews}
# 
yelp_df %>% 
    arrange(desc(review_count)) %>%
    select(-categories.title) %>% 
    distinct() %>% 
    head(30) %>% 
    mutate(name = ordered(name, levels = rev(.$name))) %>%
    ggplot(aes(y = name, x = review_count)) +
    geom_point(shape = 21, fill = "salmon", size = 2) + 
    scale_x_continuous(breaks = seq(250, 1000, 150)) +
    theme_minimal() +
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          axis.title.y = element_blank()
          ) +
    xlab("Number of Reviews") +
    ggtitle("Thirty Businesses with the Most Yelp Reviews")
```

A histogram can show us the distribution of reviews, specifically the number of 
reviews per listing. Here we can see that the bulk of the listings have less
than 100 reviews, so the businesses in the 

```{r plot_histogram_review_count_per_listing, warning=F, message=F}
yelp_df %>% 
    select(-categories.title) %>% 
    distinct() %>%
    select(review_count) %>% 
    ggplot(aes(x = review_count)) +
    geom_histogram(color = "black", fill = "#2471A3", bins = 16) + 
    scale_x_log10(breaks = c(1, 10, seq(100, 1000, 300))) + 
    scale_y_continuous(breaks = seq(0, 2500, 500)) +
    theme_minimal() + 
    theme(panel.grid.minor.x = element_blank()
          ) + 
    ggtitle("Histogram of Review Count per Yelp Listing") + 
    xlab("Review Count (Logarithmic)")
    
```

```{r}
review_mean_df <- yelp_df %>% 
    group_by(categories.title) %>% 
    count() %>% 
    rename(Listings = n) %>% 
    full_join(yelp_df %>%
                  group_by(categories.title) %>% 
                  summarise(review_mean = mean(review_count)),
              by = "categories.title"
              )

review_mean_df$bin <- cut(review_mean_df$Listings,
                          breaks = seq(0, 1700, 100),
                          labels = as.character(seq(100, 1700, 100))
                              )


review_mean_less_100_df <- review_mean_df %>% 
    filter(Listings < 101)

review_mean_less_100_df$bin <- cut(review_mean_less_100_df$Listings,
                                   breaks = seq(0, 100, 5),
                                   labels = as.character(seq(5, 100, 5))
                                   )

ggplot(review_mean_df, aes(x = bin, y = review_mean)) +
    geom_boxplot()

ggplot(review_mean_less_100_df, aes(x = bin, y = review_mean)) +
    geom_boxplot()

review_mean_less_100_df %>% 
    filter(bin == "5", review_mean > 35) %>% 
    arrange(desc(review_mean))

# yelp_df %>%
#     filter(categories.title == "Sri Lankan") %>% 
#     select(name, review_count)
```









<br><br>

#### Berliner's Love Value.



```{r number_of_listings_in_each_price_class, warning=F, message=F}
yelp_df %>% 
    select(-categories.title) %>% 
    distinct() %>% 
    select(price) %>% 
    filter(!is.na(.)) %>%
    group_by(price) %>% 
    count() %>% 
    filter(price != "$$") %>% 
    ungroup() %>% 
    mutate(price = factor(.$price,
                          ordered = T,
                          levels = c("€", "€€", "€€€", "€€€€")
                          )
           ) %>% 
    ggplot(aes(x = price, y = n, fill = price)) +
    geom_bar(stat = "identity", color = "black") + 
    theme_minimal() + 
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          legend.position = "none"
          ) + 
    ggtitle("Number of Listings in Each Price Class") + 
    xlab("Price Class") + 
    ylab("Count")
```


## Deciding on a Metric

Ilia Blinderman and Amber Thomas [draw](https://pudding.cool/2018/03/neighborhoods/)
on Katie Hempenius's reasoning in her 
[analysis](https://katiehempenius.com/post/yelp-map-of-san-francisco/) of San
Francisco's neighborhoods. With the dual goals of "identify[ing] interesting 
trends" and "normaliz[ing] for the effects of commercial versus residential 
usage" in mind, Katie arrives at the following formula for "incidence rate":

$$
X = \frac{Category\;Z\;in\;zip}{Category\;Z\;in\;city} \times
\frac{Businesses\;in\;city}{Businesses\;in\;zip}
$$

I have a number of problems with this metric, ranging from major to minor.


Luckily, one of biggest issues is also the easiest to address. Simply put, the
above formula does not describe the _incidence rate_. The [incidence rate](https://www.cdc.gov/ophss/csels/dsepd/ss1978/lesson3/section2.html) is a
measure of _new cases_ in a specified period of _time_. Time, however, doesn't
play a role in the above formula - nor should it.

This doesn't at all invalidate Katie's, Ilia's, and Amber's respective work, as 
"incidence" is commonly used to indicate the _occurence_ of something. In fact,
most dictionary [definitions](https://www.dictionary.com/browse/incidence) 
reflect just this usage. Nevertheless, the _incidence rate_ has a more specific 
meaning than _incidence_ alone, and in the present context, where an
epidemiological approach is appropriate, I think it's important to split a
few hairs.

A less pressing issue, but perhaps relevant in terms of computational resources,
is the second half of the equation. It is a constant that scales all
neighborhood-level values upwards by the same amount, and thus has no impact on
the neighborhood-level rankings, which is what we are interested in. A 
simplified formula will produce the same results:

$$
X = \frac{Category\;Z\;in\;zip}{Category\;Z\;in\;city}
$$

What remains is a simple formula for calculating the proportion of city-wide
category members that can be found in a given neighborhood. Although valuable,
to a degree, I don't believe this is the best metric for determining a 
neighborhood's character, for the following reasons: 

1. _It ignores differences in frequency_. If 10 of 20 city-wide members of 
   category c~1~ are found in a neighborhood, and 100 of 200 city-wide member of
   category c~2~ are found in the same neighborhood, then the formula tells us 
   that the neighborhood is equally characterized by c~1~ and c~2~ (10/20 = 0.5
   and 100/200 = 0.5). Even worse, if c~2~ has 210 instead of 200 city-wide 
   members, then the formula tells us that c~1~ characterizes the neighborhood 
   _better_ than c~2~, even though c~2~ occurs 10 times more than c~1~!
2. _It is biased in favor of small sample sizes_. Different types of businesses 
   likely aren't evenly distributed across a city - my analysis wouldn't make any
   sense otherwise! In particular, a small number of niche businesses might
   more likely be clustered in one area. For example, vegan restaurants might be
   fairly rare in the city as a whole, but in one particulary trendy neighborhood
   you might find a cluster of them. As is the nature of things, "trendy" 
   neighborhoods are often also gentrifying inner-city areas with large minority
   communities. If the minority population is dispersed throughout the rest of 
   the city at a rate greater than the vegan restaurants, then the above formula 
   would suggest that the neighborhood is better characterized as "vegan" than 
   as Black, Latino, Arabic, Indian, etc. even if these communities and cultures are far
   more prominent in everyday life. Of course, it's still important to know that
   there are more vegan restaurants in one neighborhood than anywhere else, but 
   neighborhood's uniqueness should not be determined primarily by the city-wide
   scarcity of some of it's businesses.
3. _It doesn't take into account variations in the total number of businesses_.
   Two neighborhoods might both be home to 20 of 100 total category members, i.e.
   each has 20% of the whole, but neighborhood _x_ has a total of 2,000 businesses,
   while neighborhood _y_ only has 200. Should we say that the category characterizes
   both neighborhoods equally? Probably not.
4. _It doesn't necessarily reflect interest in the category_. There are two 
   main indicators of interest in the Yelp data: 1. The number of listings per
   category, and 2. the number of reviews per listing. 
   
   

Calculating the proportion of category members in a district can tell us something
about what makes a neighborhood unique

$$
score = 
prevalence \times
proportion \times
interest
$$
Each part of this formula brings its own advantages and disadvantages. For 
instance, prevalence might tell us what occurs the most, but it doesn't tell us
anything about less prevalent businesses that might still be important in some
other way, e.g. in terms of 

$$
score = 
\left(
    \frac{category_{zip}}{businesses_{zip}} - 
    \frac{category_{city}}{businesses_{city}}
\right) \times
\frac{category_{zip}}{category_{city}} \times
\frac{reviews_{mean/zip}}{reviews_{mean/city}}
$$

I'll refer to the first part as the _relative prevalence_, since it is a measure
of district-level prevalence relative to the city-level prevalence. I've chosen 
to use the relative prevalence, instead of just the district-level prevalence,
because it is a better indicator of how the district is different from the city
as a whole. It also offers a handy cut-off: Anything less than 0 is less prevalent
in the district than in the city, and anything more than 0 is more prevalent in
the district than in the city. It's then just a matter of sorting the relative 
prevalences from greatest to least to find the categories that are overrepresented
in the district.





As a virtuous programmer (I try 🤷) I'll attempt to zero in on a good solution
via edge cases, which I can formulate in a series of what-ifs:

Dropping categories with small sample sizes (Ilia and Amber set a threshold of
1% of a district's businesses) ignores the issue posed by _1._, i.e. that a 
neighborhood might be characterized by a small number of unusual businesses, or
perhaps a single, very large institution. 

I think most of the confusion here stems from the fact that specific epidemiological terminology is being used in a context where an epidemological approach is perfectly valid, but how the terminology is being used isn't quite right.


Using the proportion `category in district/category in city` can be misleading.





<br>

## Friedrichshain-Kreuzberg

```{r set_up_color_map_function}
district_color_map <- unique(sf_districts$Districts)
district_color_map <- data_frame(district = district_color_map, color = "grey")

district_name <- "Friedrichshain-Kreuzberg"

highlight_district <- function(district_name, color_map = district_color_map){
    color_map[which(color_map$district == district_name), "color"] <- "red"
    color_map
}

```

```{r highlight_friedrichshain_kreuzberg, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Friedrichshain-Kreuzberg")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```


```{r set_up_dataframe_of_prevalence_diffs, message=FALSE, warning=FALSE}

# calculate category prevalence for city
preval_city_df <- yelp_df %>% 
    group_by(categories.title) %>% 
    summarise(cat_tot_city = n()) %>% 
    mutate(bus_tot_city = length(unique(yelp_df$id)),
           preval_city = cat_tot_city/nrow(yelp_df))

# calculate category prevalence for districts
preval_dist_df <- yelp_df %>% 
    group_by(district, categories.title) %>% 
    count() %>%
    full_join(yelp_df %>%
                  group_by(district) %>%
                  count(),
              by = "district"
              ) %>% 
    rename(cat_tot_dist = n.x, bus_tot_dist = n.y) %>% 
    mutate(preval_dist = cat_tot_dist/bus_tot_dist)

#
rev_dist_df <- yelp_df %>% 
    group_by(district, categories.title) %>% 
    summarise(review_dist_sum = sum(review_count),
        review_dist_mean = mean(review_count, trim = .1)) %>% 
    full_join(yelp_df %>%
                  group_by(categories.title) %>%
                  summarise(review_city_sum = sum(review_count),
                      review_city_mean = mean(review_count, trim = .1)),
              by = "categories.title"
              ) %>% 
    mutate(review_prop = review_dist_mean/review_city_mean) %>% 
    group_by(district) %>% 
    mutate(review_dist_all = sum(review_dist_sum),
           review_dist_prop = review_dist_sum/review_dist_all
           )

# calculate category prevalence * proportion reviews for district
preval_prop_dist_df <- preval_dist_df %>% 
    full_join(preval_city_df,
              by = "categories.title"
              ) %>% 
    mutate(preval_diff = preval_dist - preval_city,
           prop_cat = cat_tot_dist/cat_tot_city
           ) %>% 
    full_join(rev_dist_df,
              by = c("district", "categories.title")
              ) %>% 
    mutate(preval_x_catprop_x_revprop = preval_diff * prop_cat * review_prop) %>%
    filter(preval_dist > 0.01 | review_dist_prop > 0.01) %>% 
    arrange(district, desc(preval_x_catprop_x_revprop))

categories <- unique(preval_prop_dist_df$categories.title)
categ_list <- vector(mode = "list", length = length(categories))
names(categ_list) <- categories

for (category in categories) {
    categ_list[[category]] <- preval_prop_dist_df %>% 
        filter(categories.title == category) %>% 
        arrange(desc(preval_diff)) %>% 
        head(4)
}

final_df <- bind_rows(categ_list) %>% 
    arrange(district, desc(preval_x_catprop_x_revprop))
```

```{r}
d <- "Friedrichshain-Kreuzberg"

prop_cat_df <- preval_prop_dist_df %>%
    filter(district == d) %>% 
    arrange(desc(prop_cat)) %>% 
    head(5) %>%
    mutate(Method = "Category Proportion")

preval_dist_df <- preval_prop_dist_df %>%
    filter(district == d) %>% 
    arrange(desc(preval_dist)) %>% 
    head(5) %>%
    mutate(Method = "Prevalence")

preval_diff_df <- preval_prop_dist_df %>%
    filter(district == d) %>% 
    arrange(desc(preval_diff)) %>% 
    head(5) %>%
    mutate(Method = "Relative Prevalence")

review_prop_df <- preval_prop_dist_df %>%
    filter(district == d) %>% 
    arrange(desc(review_prop)) %>% 
    head(5) %>%
    mutate(Method = "Review Proportion")

preval_x_catprop_x_revprop_df <- preval_prop_dist_df %>%
    filter(district == d) %>% 
    arrange(desc(preval_x_catprop_x_revprop)) %>% 
    head(5) %>%
    mutate(Method = "RelPre x CatPro x RevPro")

preval_x_catprop_x_revprop_filter_df <- final_df %>%
    filter(district == d) %>% 
    arrange(desc(preval_x_catprop_x_revprop)) %>% 
    head(5) %>%
    mutate(Method = "RP x CP x RP + Comp")

bind_rows(list(prop_cat_df, preval_dist_df, preval_diff_df, review_prop_df,
               preval_x_catprop_x_revprop_df, preval_x_catprop_x_revprop_filter_df
               )
          ) %>%
    ungroup() %>% 
    mutate(Method = factor(Method, levels = unique(.$Method))
           ) %>% 
    ggplot(aes(x = categories.title, y = cat_tot_city, fill = Method)) +
    geom_bar(stat = "identity") +
    geom_hline(yintercept = 0) +
    coord_flip() +
    facet_wrap(~Method) +
    theme_minimal() +
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1),
          axis.text.y = element_text(size = 7),
          legend.position = "none"
          ) +
    ylab("City-Wide Prevalence") +
    ggtitle("A Comparison of Metrics")
```


```{r write_plot_district_function}
plot_district <- function(district_name, coords) {
    
    dist_preval_df <- final_df %>% 
        filter(district == district_name) %>% 
        head(8)
    
    dist_df <- yelp_df %>% 
        filter(district == district_name,
               categories.title %in% dist_preval_df$categories.title
               ) %>% 
        mutate(categories.title = ordered(categories.title,
                                          levels = dist_preval_df$categories.title
                                          )
               )
    
    coords_df <- tibble(lon = coords[c("top", "bottom")],
                        lat = coords[c("right", "left")]
                        )
    
    qmplot(data = coords_df,
           x = lat,
           y = lon,
           maptype = "terrain",
           geom = "blank"
           ) +
        geom_point(data = dist_df, 
                   aes(x = coordinates.longitude,
                       y = coordinates.latitude,
                       shape = I(21),
                       fill = I("red"),
                       size = review_count,
                       alpha = .5

                       ),
                   ) +
        facet_wrap(~ categories.title, ncol = 2) +
        theme_minimal() +
        theme(axis.title = element_blank(),
              axis.text = element_blank(),
              axis.ticks = element_blank(),
              strip.text = element_text(size=12),
              legend.position = "none"
              )
}

```

```{r plot_top_6_friedrichshain_kreuzberg, warning=F, message=F, fig.height=10}

coords <- c(top = 52.5193, right = 13.4677, bottom = 52.4730, left = 13.3535)
plot_district("Friedrichshain-Kreuzberg", coords = coords)

```









<br>

## Mitte

```{r highlight_mitte, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Mitte")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```


```{r plot_top_6_mitte, warning=F, message=F, fig.height=10}

coords <- c(top = 52.5794, right = 13.4797, bottom = 52.4870, left = 13.2514)
plot_district("Mitte", coords = coords)

```











<br>

## Charlottenburg-Wilmersdorf

```{r highlight_charlottenburg_wilmersdorf, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Charlottenburg-Wilmersdorf")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```



```{r plot_top_6_charlottenburg_wilmersdorf, warning=F, message=F, fig.height=10}

coords <- c(top = 52.5542, right = 13.3781, bottom = 52.4618, left = 13.1498)
plot_district("Charlottenburg-Wilmersdorf", coords = coords)

```









<br>

## Steglitz-Zehlendorf

```{r highlight_steglitz_zehlendorf, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Steglitz-Zehlendorf")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```


```{r plot_top_6_steglitz_zehlendorf, warning=F, message=F, fig.height=10}

coords <- c(top = 52.5221, right = 13.4583, bottom = 52.3368, left = 13.0016)
plot_district("Steglitz-Zehlendorf", coords = coords)

```








<br>

## Tempelhof-Schöneberg

```{r highlight_tempelhof_schöneberg, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Tempelhof-Schöneberg")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```

```{r plot_top_6_tempelhof_schöneberg, warning=F, message=F, fig.height=10}

coords <- c(top = 52.5329, right = 13.6021, bottom = 52.3477, left = 13.1455)
plot_district("Tempelhof-Schöneberg", coords = coords)

```











<br>

## Neukölln

```{r highlight_neukoelln, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Neukölln")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```

```{r plot_top_6_neukoelln, warning=F, message=F, fig.height=10}

coords <- c(top = 52.5384, right = 13.6903, bottom = 52.3532, left = 13.2337)
plot_district("Neukölln", coords = coords)

```








<br>

## Treptow-Köpenick

```{r highlight_treptow_koepenick, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Treptow-Köpenick")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```

```{r plot_top_6_treptow_koepenick, warning=F, message=F, fig.height=10}

coords <- c(top = 52.5108, right = 13.8287, bottom = 52.3255, left = 13.3721)
plot_district("Treptow-Köpenick", coords = coords)

```








<br>

## Marzahn-Hellersdorf

```{r highlight_marzahn_hellersdorf, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Marzahn-Hellersdorf")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```

```{r plot_top_6_marzahn_hellersdorf, warning=F, message=F, fig.height=10}

coords <- c(top = 52.6147, right = 13.8160, bottom = 52.4299, left = 13.3594)
plot_district("Marzahn-Hellersdorf", coords = coords)

```









<br>

## Lichtenberg

```{r highlight_lichtenberg, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Lichtenberg")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```

```{r plot_top_6_lichtenberg, warning=F, message=F, fig.height=10}

coords <- c(top = 52.6243, right = 13.7401, bottom = 52.4395, left = 13.2835)
plot_district("Lichtenberg", coords = coords)

```







<br>

## Pankow

```{r highlight_pankow, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Pankow")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```

```{r plot_top_6_pankow, warning=F, message=F, fig.height=10}

coords <- c(top = 52.6899, right = 13.6636, bottom = 52.5054, left = 13.2069)
plot_district("Pankow", coords = coords)

```









<br>

## Reinickendorf

```{r highlight_reinikendorf, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Reinickendorf")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```

```{r plot_top_6_reinikendorf, warning=F, message=F, fig.height=10}

coords <- c(top = 52.6970, right = 13.5238, bottom = 52.5125, left = 13.0672)
plot_district("Reinickendorf", coords = coords)

```








<br>

## Spandau

```{r highlight_spandau, fig.height=1.5, fig.width=2.5}
cmap <- highlight_district("Spandau")

ggplot(sf_districts) +
    geom_sf(aes(fill = cmap$color)) +
    scale_fill_manual(values = c("grey", "red"), guide = F) +
    theme(
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )
```

```{r plot_top_6_spandau, warning=F, message=F, fig.height=10}

coords <- c(top = 52.6118, right = 13.4243, bottom = 52.4269, left = 12.9676)
plot_district("Spandau", coords = coords)

```